---
Title: London Congestion Charge Programme
Author: Vy Nguyen
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data pre-processing

```{r Load libraries}
library(nloptr)
library(stargazer)
library(lattice)
library(ggplot2)
```

```{r Import data}

# Set working directory

setwd('~/Library/CloudStorage/OneDrive-City,UniversityofLondon/term 2/SMM641 Revenue Management and Pricing/Courseworks/Problem Set 2 (28:03) (checked)')

# Import the data

data=read.csv('CongestionPricing.csv')

# Inspect data structures

head(data)

str(data)

# Get the number of data observations

N=nrow(data)
```

# Scenario 1: Sinple Price Optimisation

If the programme’s objective were solely to maximise revenue and a single congestion charge were to be applied across both peak and non-peak hours, which price would maximize the total revenue? With this price in effect, what is the total level of emissions?

To fine the optimal single price, we set a single price across all time slots and all car types to find the maximised revenue and compute the according emissions.

```{r maxWTP}

# Get the maximum WTP across 2 time slots

for (i in 1:N){data$maxWTP[i]=max(data[i,2:3])}

# Check the new column

head(data)
```

```{r Optimal Single Price Search}

# The maximum WTP in data, we can use this as the upper bound for our price search
# No need to consider a price if no one can afford it.

maxprice=max(data$maxWTP)

# Define empty array variables we will be introducing

demandSinglePrice = rep(NA,maxprice)
revenue           = rep(NA,maxprice)
averageSpeed      = rep(NA,maxprice)
emissionsPerCar   = rep(NA,maxprice)
totalEmissions    = rep(NA,maxprice)
demandLondon      = rep(NA,maxprice)
price             = c(1:maxprice)

# Find how many people willing to buy at different price levels

for (p in 1:maxprice){
    # demand at each price level
    demandSinglePrice[p] = sum(data$maxWTP>=p) # total number of TRUE value
    # convert demand to represent the population level (in thousands)
    demandLondon[p] = demandSinglePrice[p]*192/N
    # total revenue of congestion charge in London
    revenue[p]=p*demandLondon[p]*1000
}

task1a = data.frame(price, demandSinglePrice, demandLondon, 
                    averageSpeed, emissionsPerCar, totalEmissions, revenue)

# Identify values to maximise Revenue

revenueBest = max(revenue) 
priceBest = which(revenue == revenueBest)
emissionsPerCarBest = emissionsPerCar[priceBest]
totalEmissionsBest = totalEmissions[priceBest]

# Print result

print(paste("If a single congestion charge were to be charged across both peak and non-peak hours, the optimal revenue is (£)", round(revenueBest),"achieved at the price level (£)", priceBest))
```

```{r Level of Emissions}

# Set empty array

demandNonPeak_single = rep(NA,N)
demandPeak_single = rep(NA,N)

# Compute the consumer surplus and classify which hours driver will enter London

for (r in 1:N){
  surplusNonPeak_single = data[r,3] - priceBest
  surplusPeak_single = data[r,2] - priceBest
  demandNonPeak_single[r] = (surplusNonPeak_single>surplusPeak_single)*(surplusNonPeak_single>=0)
  demandPeak_single[r] = (surplusPeak_single>=surplusNonPeak_single)*(surplusPeak_single>=0)
}

# Demand of total population (in thousands)

demandNonPeak_single_total = sum(demandNonPeak_single) * (192/N)
demandPeak_single_total = sum(demandPeak_single) * (192/N)

# Compute average speed and emissions level at 2 periods

avg_speed_NonPeak = 30 - 0.0625*demandNonPeak_single_total
emissions_NonPeak = (ifelse(avg_speed_NonPeak<25,
                           617.5-16.7*avg_speed_NonPeak,
                           235.0-1.4*avg_speed_NonPeak))*demandNonPeak_single_total

avg_speed_Peak = 30 - 0.0625*demandPeak_single_total
emissions_Peak = (ifelse(avg_speed_Peak<25,
                           617.5-16.7*avg_speed_Peak,
                           235.0-1.4*avg_speed_Peak))*demandPeak_single_total

# Level of emssions all day 

emissionsLevel = emissions_NonPeak + emissions_Peak

# Print result

print(paste("At", priceBest,", the total level of emissions is",
            round(emissionsLevel),"(g/km)"))
```


```{r Plot Number of Cars vs Single Price}

# Plotting Number of cars vs Price

xaxis=1:maxprice

plot(xaxis, demandLondon,
     pch = 16, type="s", col="blue", las=1, xaxt="n",
     xlab="Congestion Charge (£)",ylab="Number of Demands (in thousands)",
     main = "Number of Demands at Different Levels of Congestion Charge")

xticks <- seq(0, maxprice, by=2)

axis(side = 1, at = xticks)
```

```{r Plot Revenue vs Single Price}

# Plotting Revenue vs Price

xaxis=1:maxprice

plot(xaxis,revenue/1000,pch = 16, type="s",col="blue",las=1, xaxt="n",
     xlab="Congestion Charge (£)",ylab="Generated Funds (in thousands £)",
     main = "Generated Funds and Congestion Charge")

xticks <- seq(0, maxprice, by=2)

axis(side = 1, at = xticks) 

axis(side = 1, at = priceBest) 

lines(c(priceBest,priceBest),c(0, revenueBest/1000),lty=2)

axis(side = 2, at = round(revenueBest/1000,3),las=1,pos=20, tick=F)

lines(c(20,priceBest),c(revenueBest/1000, revenueBest/1000),lty=2)
```

# Scenario 2: Find Optimal Peak Price at Base Price Level £7

If the programme’s objective were solely to maximise revenue and a peak period pricing
strategy is to be implemented, with the price for the non-peak period set at £7, what price would we recommend for the peak period? Note the resulting revenue and
emissions and compare the findings with those from scenario 1.

```{r Step 1: Nonpeak Maximum Surplus}

basePrice = 7

data$surplusNonPeak = data$Nonpeak_WTP-basePrice

data[1:10,]
```

```{r Step 2: Peak Maximum Surplus}

# Create a matrix of N rows, max price columns, fill with all 0 value
surplusPeak = matrix(0,N,maxprice) 

# Compute Maximum Peak Surplus

for (p in 1:maxprice){
    for (i in 1:N){
        surplusPeak[i,p]=data[i,2]-p
    }
}

# Show results

colnames(surplusPeak)=paste0("p=",1:maxprice)

surplusPeak[1:10,]
```

```{r Step 3: Demands for Peak and NonPeak hours}

# Create empty arrays

demandNonPeak<-rep(0,maxprice)

demandPeak<-rep(0,maxprice)

revenue2<-rep(0,maxprice)

emissions1b<-rep(0,maxprice)

demand1b<-rep(0,maxprice)

# Compare consumer surplus and classify customer segments into peak and non-peak hours

for (p in 1:maxprice){
  
  surplusNonPeak = data$surplusNonPeak
  
  demandNonPeak[p]=sum((surplusNonPeak>surplusPeak[,p])*(surplusNonPeak>=0)) * (192/N)
  
  demandPeak[p]=sum((surplusPeak[,p]>=surplusNonPeak)*(surplusPeak[,p]>=0)) * (192/N)
  
  revenue2[p]=basePrice*demandNonPeak[p]*1000+p*demandPeak[p]*1000
  
  # Find level of emissions at each price point
  
  ## Non Peak hours
  
  avg_speed_NonPeak = 30 - 0.0625*demandNonPeak[p]
  
  emissions_NonPeak = (ifelse(avg_speed_NonPeak<25,
                           617.5-16.7*avg_speed_NonPeak,
                           235.0-1.4*avg_speed_NonPeak))*demandNonPeak[p]
  ## Peak hours
  
  avg_speed_Peak = 30 - 0.0625*demandPeak[p]
  
  emissions_Peak = (ifelse(avg_speed_Peak<25,
                           617.5-16.7*avg_speed_Peak,
                           235.0-1.4*avg_speed_Peak))*demandPeak[p]

  # Level of emissions all day 
  
  emissions1b[p] = emissions_NonPeak + emissions_Peak
  
  emissions1b[p]
}

task1b = data.frame(peakPrice=c(1:maxprice),
                    demandNonPeak, demandPeak, 
                    revenue2, emissions1b)
task1b
```

```{r Maximum Revenue}

# Identify values to maximise Revenue

revenueBest2 = max(revenue2)

priceBest2 = which(revenue2 == revenueBest2)

emissionsBest2 = emissions1b[priceBest2]

# Print result

print(paste("When Non-peak periods have a base price of 7 (£), the optimal revenue is", round(revenueBest2), "(£) at the optimal price for the Peak-hours slot is", priceBest2, "(£)"))

print(paste("Level of emssisions for this price point is", round(emissionsBest2), "(g/km)"))
```

```{r Plot Non-Peak Demand vs Peak Period Price}

# Plotting NonPeak Demand vs Peak Period Price

xaxis=1:maxprice

plot(xaxis,task1b$demandNonPeak,pch = 16, type="s",col="blue", las=1, xaxt="n",
     xlab="Congestion Charge  for Peak Period (£)",ylab="Non-Peak Demand (in thousands)",
     main = "NonPeak Demand vs Peak Period Congestion Charge ")

xticks <- seq(0, maxprice, by=2)

axis(side = 1, at = xticks)
```


```{r Plotting Peak Demand vs Peak Period Price}

# Plotting Peak Demand vs Peak Period Price

xaxis=1:maxprice

plot(xaxis,task1b$demandPeak,pch = 16, type="s",col="blue", las=1, xaxt="n",
     xlab="Congestion Charge for Peak Period (£)",ylab="Peak Period Demand (in thousands)",
     main = "Peak Demand vs Peak Period Congestion Charge")

xticks <- seq(0, maxprice, by=2)

axis(side = 1, at = xticks)
```

```{r Plotting Revenue vs Peak Period Price}

# Plotting Revenue vs Peak Period Price

xaxis=1:maxprice

plot(xaxis,task1b$revenue2/1000,pch = 16, type="s",col="blue", las=1, xaxt="n",
     xlab="Price for Peak Period",ylab="Generated Funds (thousands £)",
     main = "Generated Funds vs Peak Period Price")

xticks <- seq(0, maxprice, by=2)

axis(side = 1, at = xticks)

revenueBest2 = max(revenue2[basePrice:maxprice])

priceBest2 = which(revenue2/1000 == revenueBest2/1000)

axis(side = 1, at = priceBest2) 

lines(c(priceBest2,priceBest2),c(0, revenueBest2),lty=2)

axis(side = 4, at = round(revenueBest2/1000,3),las=1, pos=17.1, tick=F)

lines(c(priceBest2,20),c(revenueBest2/1000, revenueBest2/1000),lty=2)
```

# Scenrio 3: Optimisation  

Suppose now that the programme’s objective is to minimize emissions rather than
maximize revenue. However, the City would like to ensure that the programme can
self-sustain its operation and that a sufficient portion of the revenue is allocated to reinvest in the public transportation infrastructure. Overall, the City requires that the revenue should not fall below £1.1 million per day. 

Assuming a non-peak period price of £7, what price would we recommend for the peak period? Compare the resulting revenue and emissions level with that of part of Scenario 2.

```{r Fit Linear Models}

# Fit linear model for multiple price

fitNonPeak = lm(demandNonPeak~peakPrice, data=task1b)

a1=coef(fitNonPeak)[1]

b1=coef(fitNonPeak)[2]

summary(fitNonPeak)

fitPeak = lm(demandPeak~peakPrice, data=task1b)

a2=coef(fitPeak)[1]

b2=coef(fitPeak)[2]

summary(fitPeak)
```

The results show a great fit to the data points with significantly high R-squared.

```{r Use Linear Models to Solve Optimisation Problem}

demandNonPeak=rep(0,maxprice)

demandPeak=rep(0,maxprice)

demandTotal=rep(0,maxprice)

emissionsNonPeak=rep(0,maxprice)

emissionsPeak=rep(0,maxprice)

emissionsTotal=rep(0,maxprice)

revenue1c=rep(0,maxprice)

for (p in (1:maxprice)) {
  
  demandNonPeak[p]=a1+b1*p
  
  demandPeak[p]=a2+b2*p
  
  demandTotal[p]=demandNonPeak[p]+demandPeak[p] # in thousand
  
  averageSpeed_NonPeak=30-0.0624*demandNonPeak[p]

  emissionsNonPeak[p] = (ifelse(avg_speed_NonPeak<25,
                           617.5-16.7*avg_speed_NonPeak,
                           235.0-1.4*avg_speed_NonPeak))*demandNonPeak[p]
  
  averageSpeed_Peak=30-0.0624*demandPeak[p]
  
  emissionsPeak[p] = (ifelse(averageSpeed_Peak<25,
                           617.5-16.7*averageSpeed_Peak,
                           235.0-1.4*averageSpeed_Peak))*demandPeak[p]
  
  emissionsTotal[p]=emissionsNonPeak[p] + emissionsPeak[p]
  
  revenue1c[p]=7*demandNonPeak[p]+p*demandPeak[p]
}

task1c = data.frame(c(1:17), demandNonPeak, demandPeak, demandTotal, 
                    emissionsNonPeak, emissionsPeak,
                    emissionsTotal, revenue1c) 
task1c
```

```{r Classify revenue on constraint}

task1c=task1c[task1c$revenue1c>=1100,]
task1c
```

```{r Identify the optimal metrics}

# Identify the optimal result

emissionsBest1c = min(task1c$emissionsTotal)

priceBest1c = which(emissionsTotal == emissionsBest1c)

revenueBest1c = revenue1c[priceBest1c] # in thousand

print(paste("Minimum emissions level of (g/km)", round(emissionsBest1c), "is achieved at price (£)", round(priceBest1c), "with the generated refund of (£)", round(revenueBest1c*1000))
)
```

